---
title: "Retractiowatch project: Preregistered analysis"
author: "Marton Kovacs"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
```

# Load data

_Note:_ The name of the datafile will be changed only for the final analysis.

```{r}
processed <- read_csv(here("data/processed/retractionwatch_processed_simulated_data.csv"))
```

# Distribution of completition time

On median it took `r median(processed$duration_in_seconds) / 60` to finish the survey to the participants in our sample.

```{r}
# We will only need the duration in minutes for the main analysis so we calculate it here
processed <-
  processed %>% 
  mutate(duration_in_mins = duration_in_seconds / 60)

processed %>% 
  ggplot() +
  aes(x = duration_in_mins) +
  geom_histogram() +
  labs(
    x = "Response duration in minutes",
    y = "Count"
  )
```

# To what degree was the retraction of your work, mentioned in our email, caused by data management mistakes?

```{r}
# With percentage
mistake_reason_plot_data <- 
  processed %>%
  select(reason_mistake) %>% 
  mutate(reason_mistake = factor(reason_mistake, levels = c(
        "It was the main reason",
        "Partially",
        "Not at all"
      ))) %>% 
  count(reason_mistake) %>% 
  ungroup() %>% 
  mutate(N = sum(n),
         percentage = round(n / N * 100, 0))

y_max <- max(mistake_reason_plot_data$n) + 10
  
mistake_reason_plot <- 
  mistake_reason_plot_data %>% 
  ggplot() +
  aes(x = reason_mistake, y = n) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = paste0(percentage, "%")), size = 8, vjust = -0.3) +
  labs(x = paste0("To what degree was the retraction of your work,\nmentioned in our email, caused by data management mistakes? (n =", unique(mistake_reason_plot_data$N), ")"),
       y = "Count") +
  scale_y_continuous(expand = c(0,0), limit = c(0, y_max)) +
  theme(axis.ticks = element_line(color = "black"),
        axis.line = element_line(color = "black"),
        axis.text = element_text(color = "black", size = 18),
        panel.background = element_blank(),
        axis.title = element_text(size = 18),
        plot.margin = unit(c(0,0.8,0,0), "cm"))

mistake_reason_plot
```

# Which of the following data management mistake type contributed the most to the retraction of your paper?

We calculate the percentage of occurrence for all RDM mistake types. We also create a figure on the distribution of these percentages for those mistake types that occurred in more than 1% of the cases.

```{r}
mistake_type_plot_data <-
  processed %>% 
  # Excluding irrelevant responses
  filter(mistake_type != "Irrelevant") %>% 
  select(mistake_type) %>% 
  count(mistake_type) %>% 
  ungroup() %>% 
  mutate(N = sum(n),
         percentage = round(n / N * 100, 0)) %>% 
  arrange(desc(n))

y_max <- max(mistake_type_plot_data$n) + 10
  
mistake_type_plot <-
  mistake_type_plot_data %>% 
  filter(n > 1) %>% 
  ggplot() +
  aes(x = reorder(mistake_type, n), y = n) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = paste0(percentage, "%")), size = 8, hjust = -0.3) +
  labs(x = paste0("Which of the following data management mistake type\ncontributed the most to the retraction of your paper? (n =", unique(mistake_type_plot_data$N), ")"),
       y = "Count") +
  coord_flip() +
  scale_y_continuous(expand = c(0,0), limit = c(0, y_max)) +
  theme(
    axis.ticks = element_line(color = "black"),
    axis.line = element_line(color = "black"),
    axis.text = element_text(color = "black", size = 18),
    panel.background = element_blank(),
    axis.title = element_text(size = 18),
    plot.margin = unit(c(0,0.8,0,0), "cm")
    )

mistake_type_plot
```

The number of irrelevant responses.

```{r}
processed %>% 
  filter(mistake_type == "Irrelevant") %>%
  nrow()
```

This is the figure for all the responses. This will be presented in the supplementary materials.

```{r}
mistake_type_all_plot <-
  mistake_type_plot_data %>% 
  ggplot() +
  aes(x = reorder(mistake_type, n), y = n) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = paste0(percentage, "%")), size = 8, hjust = -0.3) +
  labs(x = paste0("Which of the following data management mistake type\ncontributed the most to the retraction of your paper? (n =", unique(mistake_type_plot_data$N), ")"),
       y = "Count") +
  coord_flip() +
  scale_y_continuous(expand = c(0,0), limit = c(0, y_max)) +
  theme(
    axis.ticks = element_line(color = "black"),
    axis.line = element_line(color = "black"),
    axis.text = element_text(color = "black", size = 18),
    panel.background = element_blank(),
    axis.title = element_text(size = 18),
    plot.margin = unit(c(0,0.8,0,0), "cm")
    )

mistake_type_all_plot
```

# What was the main cause of the data management mistake you identified above? 

```{r}
mistake_cause_plot_data <-
  processed %>% 
  select(mistake_cause) %>% 
  # Excluding irrelevant responses
  filter(mistake_cause != "Irrelevant") %>% 
  count(mistake_cause) %>% 
  ungroup() %>% 
  mutate(N = sum(n),
         percentage = round(n / N * 100, 0)) %>% 
  arrange(desc(n))

y_max <- max(mistake_cause_plot_data$n) * 1.2
  
mistake_cause_plot <-
  mistake_cause_plot_data %>% 
  filter(n > 1) %>% 
  ggplot() +
  aes(x = reorder(mistake_cause, n), y = n) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = paste0(percentage, "%")), size = 8, hjust = -0.3) +
  labs(x = paste0("What was the main cause of the\ndata management mistake you identified above?  (n =", unique(mistake_cause_plot_data$N), ")"),
       y = "Count") +
  coord_flip() +
  scale_y_continuous(expand = c(0,0), limit = c(0, y_max)) +
  theme(
    axis.ticks = element_line(color = "black"),
    axis.line = element_line(color = "black"),
    axis.text = element_text(color = "black", size = 18),
    panel.background = element_blank(),
    axis.title = element_text(size = 18),
    plot.margin = unit(c(0,0.8,0,0), "cm")
    )

mistake_cause_plot
```

The number of irrelevant responses.

```{r}
processed %>% 
  filter(mistake_cause == "Irrelevant") %>%
  nrow()
```

The plot including all the responses for the supplementary materials.

```{r}
mistake_cause_plot <-
  mistake_cause_plot_data %>% 
  ggplot() +
  aes(x = reorder(mistake_cause, n), y = n) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = paste0(percentage, "%")), size = 8, hjust = -0.3) +
  labs(x = paste0("What was the main cause of the\ndata management mistake you identified above?  (n =", unique(mistake_cause_plot_data$N), ")"),
       y = "Count") +
  coord_flip() +
  scale_y_continuous(expand = c(0,0), limit = c(0, y_max)) +
  theme(
    axis.ticks = element_line(color = "black"),
    axis.line = element_line(color = "black"),
    axis.text = element_text(color = "black", size = 18),
    panel.background = element_blank(),
    axis.title = element_text(size = 18),
    plot.margin = unit(c(0,0.8,0,0), "cm")
    )

mistake_cause_plot
```

# How much stress did you experience due to the retraction of the article?

```{r}
stress_plot_data <-
  processed %>% 
  select(stress) %>% 
  count(stress) %>% 
  ungroup() %>% 
  mutate(
    N = sum(n),
    percentage = round(n / N * 100, 0),
    stress = case_when(
      stress == 0 ~ "No stress at all 0",
      stress == 1 ~ "1",
      stress == 2 ~ "2",
      stress == 3 ~ "3",
      stress == 4 ~ "4",
      stress == 5 ~ "5",
      stress == 6 ~ "6 Extreme stress"
      ),
    stress = factor(stress, levels = c("No stress at all 0" , "1" , "2" , "3" , "4" , "5" , "6 Extreme stress"))
    )

y_max <- max(stress_plot_data$n) * 1.2

stress_plot <-
  stress_plot_data %>% 
  ggplot() +
  aes(x = stress, y = n) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = paste0(percentage, "%")), size = 8, vjust = -0.3) +
  labs(x = paste0("How much stress did you experience due to the retraction of the article? (n =", unique(stress_plot_data$N), ")"),
       y = "Count") +
  scale_y_continuous(expand = c(0,0), limit = c(0, y_max)) +
  theme(
    axis.ticks = element_line(color = "black"),
    axis.line = element_line(color = "black"),
    axis.text = element_text(color = "black", size = 18),
    panel.background = element_blank(),
    axis.title = element_text(size = 18),
    plot.margin = unit(c(0,0.8,0,0), "cm")
    )

stress_plot
```

We will calculate the median and IQR for the question.

```{r}
# Median
median(processed$stress, na.rm = TRUE)

# Quantile 25%
quantile(processed$stress, .25, na.rm = TRUE)

# Quantile 75%
quantile(processed$stress, .27, na.rm = TRUE)
```

# Did you change anything in your research workflow because of the retraction?

```{r}
workflow_plot_data <-
  processed %>% 
  select(workflow) %>% 
  count(workflow) %>% 
  ungroup() %>% 
  mutate(
    N = sum(n),
    percentage = round(n / N * 100, 0)
    )

y_max <- max(workflow_plot_data$n) * 1.2

workflow_plot <-
  workflow_plot_data %>% 
  ggplot() +
  aes(x = workflow, y = n) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = paste0(percentage, "%")), size = 8, vjust = -0.3) +
  labs(x = paste0("Did you change anything in your research workflow because of the retraction? (n =", unique(workflow_plot_data$N), ")"),
       y = "Count") +
  scale_y_continuous(expand = c(0,0), limit = c(0, y_max)) +
  theme(
    axis.ticks = element_line(color = "black"),
    axis.line = element_line(color = "black"),
    axis.text = element_text(color = "black", size = 18),
    panel.background = element_blank(),
    axis.title = element_text(size = 18),
    plot.margin = unit(c(0,0.8,0,0), "cm")
    )

workflow_plot
```
