---
title: "Retractiowatch project: selecting contact sample"
author: "Marton Kovacs"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
```

Loading the full Retraction Watch database provided by the Retraction Watch admins. We cannot share the database, but the following code describes our method for selecting the contact database.

# Load data

```{r}
rw_database <- readxl::read_xlsx(here::here("data/source/RWDBDNLD11102022.xlsx"))
```

# Filter retracted papers due to RMD mistakes

```{r}
# Preparing terms to filter for
filter_terms <- "Concerns/Issues About Data|Concerns/Issues About Results|Duplication of Data|Error in Analyses|Error in Data|Error in Results and/or Conclusions|Results Not Reproducible|Unreliable Data|Unreliable Results"

# Filtering
rw_filtered <-
  rw_database %>% 
  filter(str_detect(Reason, filter_terms))

# Number of hits
nrow(rw_filtered)

# Number of papers when the retraction notice is available
rw_filtered %>% 
  mutate(retraction_doi_available = case_when(
    RetractionDOI == "Unavailable" ~ 0L,
    RetractionDOI == "unavailable" ~ 0L,
    is.na(RetractionDOI) ~ 0L,
    TRUE ~ 1L
    )
  ) %>% 
  count(retraction_doi_available)

# Check in how many instances the DOI of the original paper is available
rw_filtered %>% 
  mutate(original_doi_available = case_when(
    OriginalPaperDOI == "Unavailable" ~ 0L,
    OriginalPaperDOI == "unavailable" ~ 0L,
    is.na(OriginalPaperDOI) ~ 0L,
    TRUE ~ 1L
    )
  ) %>% 
  count(original_doi_available)

# Check in how many instances the PubMed Id of the original paper is available
rw_filtered %>% 
  mutate(original_pubmedid_available = case_when(
    OriginalPaperPubMedID == 0 ~ 0L,
    is.na(OriginalPaperPubMedID) ~ 0L,
    TRUE ~ 1L
    )
  ) %>% 
  count(original_pubmedid_available)
```

# Search for duplicates

It is possible that a paper was retracted multiple times. However, we do not want to contact the same authors multiple times.

```{r}
rw_filtered_duplicates <- 
  rw_filtered %>% 
  group_by(Title) %>% 
  mutate(n = n()) %>% 
  filter(n > 1)
```


