---
title: "Retractiowatch project: Preprocessing of raw data"
author: "Marton Kovacs"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(writexl)

# Source R scripts
r_scripts <- list.files(here::here("R/"), full.names = TRUE)
purrr::walk(r_scripts, source)
```

# Load data

For the analysis the name of the raw datafile will be changed. For the preregistration we use the simulated datafile.

```{r}
raw <- read_csv(here("data/raw/retractionwatch_raw_simulated_data.csv"))
```

Before any exclusions we had `r nrow(raw)` responses in the dataset.

Qualtrics records two additional rows with metadata. We do not need these rows so we will drop them.

```{r}
raw <-
  raw %>% 
  slice(-(1:2))
```

# Exclusions
## Test data

The survey was sent out on _sent_out_date_ to the participants. Thus, any responses in the dataset before that date are coming from the testing of the survey. These responses will be excluded from further analysis. We excluded `r raw %>% dplyr::filter(lubridate::as_date(start_date) < lubridate::as_date("2023-01-28")) %>% nrow()` test responses. 

```{r}
# The current start date is just a random value for the sake of the preregistration. It will be replaced for the analysis with the actual sent out date.
raw <-
  raw %>% 
  dplyr::filter(lubridate::as_date(start_date) >= lubridate::as_date("2023-01-03"))
```

## Responses without consent

There were `r filter(raw, consent != "Yes") %>% nrow()` that did no provide consent for participation.

```{r}
raw <-
  raw %>% 
  filter(consent == "Yes")
```

## Partial responses

We only include full responses in the final analysis as preregistered. A response is considered complete if all the required questions are answered (variables _reason_mistake_, _mistake_type_, _mistake_cause_, and _stress_).

```{r}
raw <- 
  raw %>% 
  mutate(
    has_missing = case_when(
      if_any(
        c(
          reason_mistake,
          mistake_type,
          mistake_cause,
          stress
          ),
        .fns = is.na) ~ TRUE,
      TRUE ~ FALSE
      )
    )
```

There were `r dplyr::filter(raw, has_missing) %>% nrow()` respondents who did not finish the survey.

```{r}
raw <-
  raw %>% 
  filter(!has_missing)
```

At the end of the exclusion we had `r nrow(raw)` responses remaining for the analysis.

# Delete not used variables

```{r}
raw <-
  raw %>% 
  select(
    # variables created by Qualtrics that we do not need anymore
    - start_date,
    - end_date,
    - status,
    - progress,
    - finished,
    - recorded_date,
    - distribution_channel,
    - user_language,
    # Missing responses flag variables
    - has_missing
  )
```

# Grouping free-text responses

We will use thematic analysis to group the free-text responses separately for variables `mistake_type_21_text`, `mistake_cause_16_text`, `experience`.

One author will first go through the free-text responses and summarize their content into short text codes.

```{r}
to_code <-
  raw %>% 
  select(response_id, mistake_type_21_text, mistake_cause_16_text, experience) %>% 
  pivot_longer(-response_id, names_to = "variable", values_to = "value") %>% 
  filter(!is.na(value)) %>%
  arrange(variable) %>% 
  mutate(code = NA_character_)

# Danger: By running this code you can override manually coded values
# write_xlsx(to_code, here("data/raw/grouping/retractionwatch_coding.xlsx"))
```

Reading the codes and dropping irrelevant responses.

```{r}
codes <- readxl::read_xlsx(here("data/raw/grouping/retractionwatch_coding.xlsx"))

to_group <-
  codes %>% 
  filter(code != "irrelevant") %>% 
  select(- value) %>% 
  mutate(group = NA_character_)

# Danger: By running this code you can override manually coded values
# write_xlsx(to_group, here("data/raw/grouping/retractionwatch_grouping.xlsx"))
```

Creating definitions for the groups based on the codes that were assigned to them.

```{r}
groups <- readxl::read_xlsx(here("data/raw/grouping/retractionwatch_grouping.xlsx"))

definitions <-
  groups %>% 
  group_by(variable, group) %>% 
  distinct(code, .keep_all = TRUE) %>% 
  dplyr::summarise(definition = stringr::str_c(code, collapse = "; ")) %>% 
  mutate(definition = tolower(definition))

# write_xlsx(definitions, here("data/raw/grouping/retractionwatch_definitions.xlsx"))
```

Preparing data for the sorting free-text responses into groups.

```{r}
to_sort <-
  raw %>% 
  select(response_id, mistake_type_21_text, mistake_cause_16_text, experience) %>% 
  pivot_longer(-response_id, names_to = "variable", values_to = "value") %>% 
  filter(!is.na(value)) %>%
  arrange(variable) %>% 
  mutate(
    rater_1 = NA_character_,
    rater_2 = NA_character_)

# Danger: By running this code you can override manually coded values
# write_xlsx(to_sort, here("data/raw/grouping/retractionwatch_sorting.xlsx"))
```

Comparing ratings.

```{r}
sorting <- readxl::read_xlsx(here("data/raw/grouping/retractionwatch_sorting.xlsx"))

to_match <-
  sorting %>% 
  mutate(
    match = case_when(
      rater_1 == rater_2 ~ TRUE,
      TRUE ~ FALSE
    ),
    final_group = case_when(
      match == TRUE ~ rater_1,
      TRUE ~ NA_character_
    )
  )

# Danger: By running this code you can override manually coded values
# write_xlsx(to_match, here("data/raw/grouping/retractionwatch_comparison.xlsx"))
```

Merging the final groups with the raw data.

```{r}
comparison <- readxl::read_xlsx(here("data/raw/grouping/retractionwatch_comparison.xlsx"))

comparison <-
  comparison %>% 
  select(
    response_id,
    variable,
    final_group
  ) %>% 
  pivot_wider(
    id_cols = response_id,
    names_from = variable,
    values_from = final_group,
    names_prefix = "group_"
  )

raw <-
  raw %>% 
  left_join(., comparison, by = "response_id") %>% 
  mutate(
    mistake_type = case_when(
      mistake_type == "Other, please specify:" ~ group_mistake_type_21_text,
      TRUE ~ mistake_type
    ),
    mistake_cause = case_when(
      mistake_cause == "Other, please specify:" ~ group_mistake_cause_16_text,
      TRUE ~ mistake_cause
    )
  ) %>% 
  select(
    - experience,
    - group_mistake_type_21_text,
    - group_mistake_cause_16_text,
    - mistake_type_21_text,
    - mistake_cause_16_text,
    - mistake_type_21_text_modified,
    - mistake_cause_16_text_modified,
    - experience_modified
    ) %>% 
  rename(experience = group_experience)
```

# Mutating variables

In order to summarize the data in some variables in a concise manner, we clean the values in these variables.

```{r}
raw <-
  raw %>% 
  mutate(
    mistake_type = str_remove(mistake_type, "\\s*\\(.*"),
    mistake_type =  str_to_sentence(mistake_type),
    mistake_cause = str_remove(mistake_cause, "\\s*\\(.*"),
    mistake_cause =  str_to_sentence(mistake_cause),
    stress = as.integer(str_extract(stress, "\\d+")),
    reason_mistake = str_to_sentence(reason_mistake),
    experience = str_to_sentence(experience)
  )
```

# Saving the processed data

The file name of the actual datafile will be changed to _retractionwatch_processed_data.tsv_. We do not make any other modifications to the datafile.

```{r}
write_csv(raw, here("data/processed/retractionwatch_processed_simulated_data.csv"))
```