---
title: "Retractiowatch project: Preprocessing of source data"
author: "Marton Kovacs"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
```

# Load data

```{r}
source <- read_csv(here("data/source/ignore/qualtrics/retractionwatch_survey_July+25,+2023_11.40.csv"))
```

# Clean variable names

We follow the snake-case standard naming convention.

```{r}
source <- janitor::clean_names(source)
```

# Saving non-data management mistakes

To explore the retraction notices of those respondents who claim their paper was not retracted due to data management mistakes we save their responses with the title of their paper in a separate file. In order to ensure anonymity we cannot share openly the title of their paper.

```{r}
rw_database <- readxl::read_xlsx(here::here("data/source/ignore/RWDBDNLD11102022.xlsx")) %>% janitor::clean_names()

source %>% 
  dplyr::filter(
    # Title can be empty for test responses only
    !is.na(title),
    reason_mistake == "not at all"
    ) %>% 
  select(response_id, reason_mistake, title) %>% 
  # We are joining the rest of the retraction watch database to our respondents to find links easily to
  # the retraction notices
  # Note: In some cases the join will not work since the rw_database titles were not unicode and we renamed them
  left_join(., rw_database, by = "title") %>%
  mutate(is_cause_rdm = NA_character_) %>% 
  writexl::write_xlsx(., here("data/source/ignore/retractionwatch_non-mistake_data.xlsx"))
```

# Reviewing free-text variables

It is possible that to questions with free-text response options the respondents included information in their response that would make their identity compromised. We review these variables and delete the part of their response that would allow others to identify their response. These responses will be flagged.

The variables that have free-text response options are the following:
* mistake_type_21_text
* mistake_cause_16_text
* experience
* comment

We originally wanted to delete the __comment__ variable from the raw data since we planned to ask only about the respondents willingness to do a followup qualitative interview regarding their experiences with the retraction. Since some of the respondents provided useful information in this variable we decided to delete the identifying information from the variable but to keep the rest of their responses. Rows where identifying information was deleted are flagged.

There were some identifying information in the __experience__ variable as well. This variable will be modified as well similarly to the __comment__ variable.

Deleting the identifying information was done manually from the raw dataset after it has been saved.

```{r}
source <-
  source %>% 
  mutate(
    comment_modified = case_when(
      response_id == "R_2eRNkGLXhavS7Jm" ~ "modified",
      response_id == "R_2z7ujRi4un2YN96" ~ "modified",
      response_id == "R_2CHB420Ko7fYH24" ~ "modified",
      response_id == "R_3PHWZMgHssVJ0ea" ~ "modified",
      response_id == "R_1Quf6sG5SUnjG8T" ~ "modified",
      response_id == "R_3oHqfVjUBTssSki" ~ "modified",
      response_id == "R_1K7SMT7IR2f7NaX" ~ "modified",
      response_id == "R_Y02jyStUzz2j1LP" ~ "modified",
      TRUE ~ "not modified"
    ),
    experience_modified = case_when(
      response_id == "R_Y02jyStUzz2j1LP" ~ "modified",
      TRUE ~ "not modified"
    ),
  )
```

However, we do delete the title of their study from the dataset which was collected as embedded data in Qualtrics.

```{r}
source <-
  source %>% 
  select(
    -title
    )
```

# Saving the raw data

_Warning:_ The __experience__ and __comment__ variable responses that included identifiable information are removed by hand after the raw datafile has been saved. Saving the raw datafile again might override the manually modified values!

The file name of the actual datafile will be changed to _retractionwatch_raw_data.tsv_. We do not make any other modifications to the datafile.

```{r}
write_csv(source, here("data/raw/retractionwatch_raw_data.csv"))
```

