---
title: "Retractiowatch project: Preregistered analysis"
author: "Marton Kovacs"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(ggalluvial)
```

# Load data

```{r}
processed <- read_csv(here("data/processed/retractionwatch_processed_data.csv"))
```

# Distribution of completition time

On median it took `r median(processed$duration_in_seconds) / 60` to finish the survey to the participants in our sample.

```{r}
# We will only need the duration in minutes for the main analysis so we calculate it here
processed <-
  processed |>  
  mutate(duration_in_mins = duration_in_seconds / 60)

processed |> 
  ggplot() +
  aes(x = duration_in_mins) +
  geom_histogram() +
  labs(
    x = "Response duration in minutes",
    y = "Number of responses"
  )
```

# Which of the following data management error type contributed the most to the retraction of your paper?

We calculate the percentage of occurrence for all RDM error types. We also create a figure on the distribution of these percentages for those error types that occurred in more than 1% of the cases.

```{r typePlot}
mistake_type_plot_data <-
  processed |> 
  # Excluding missing and dropped responses
  filter(!is.na(mistake_type)) |> 
  select(mistake_type) |> 
  count(mistake_type) |> 
  ungroup() |> 
  mutate(N = sum(n),
         percentage = round(n / N * 100, 0)) |> 
  arrange(desc(n))

glue::glue_collapse(mistake_type_plot_data$mistake_type, sep = "\n")

# y_max <- max(mistake_type_plot_data$n) + 10
  
mistake_type_plot <-
  mistake_type_plot_data |> 
  filter(n > 1) |> 
  ggplot() +
  aes(x = reorder(mistake_type, n), y = n) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = paste0(percentage, "%")), size = 8, hjust = -0.3) +
  labs(x = paste0("Which of the following data management mistake type\ncontributed the most to the retraction of your paper? (n =", unique(mistake_type_plot_data$N), ")"),
       y = "Number of responses") +
  coord_flip() +
  scale_y_continuous(expand = c(0,0), limit = c(0, 20)) +
  theme(
    axis.ticks = element_line(color = "black"),
    axis.line = element_line(color = "black"),
    axis.text = element_text(color = "black", size = 15),
    panel.background = element_blank(),
    axis.title = element_text(size = 20),
    plot.margin = unit(c(0.5, 0.5, 0.5, 0.5), "cm")
    )

ggsave(here::here("figures/mistake_type_plot.jpg"), mistake_type_plot, dpi = 300, height = 10, width = 14)

mistake_type_plot
```

The number of missing or dropped responses.

```{r}
processed |> 
  filter(is.na(mistake_type)) |>
  nrow()
```

# What was the main cause of the data management mistake you identified above? 

```{r causePlot}
mistake_cause_plot_data <-
  processed |> 
  select(mistake_cause) |> 
  # Excluding missing and irrelevant responses
  filter(!is.na(mistake_cause)) |> 
  count(mistake_cause) |> 
  ungroup() |> 
  mutate(N = sum(n),
         percentage = round(n / N * 100, 0)) |> 
  arrange(desc(n))

# y_max <- max(mistake_cause_plot_data$n) * 1.2
  
mistake_cause_plot <-
  mistake_cause_plot_data |> 
  filter(n > 1) |> 
  ggplot() +
  aes(x = reorder(mistake_cause, n), y = n) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = paste0(percentage, "%")), size = 8, hjust = -0.3) +
  labs(x = paste0("What was the main cause of the\ndata management mistake you identified above?  (n =", unique(mistake_cause_plot_data$N), ")"),
       y = "Number of responses") +
  coord_flip() +
  scale_y_continuous(expand = c(0,0), limit = c(0, 20)) +
  theme(
    axis.ticks = element_line(color = "black"),
    axis.line = element_line(color = "black"),
    axis.text = element_text(color = "black", size = 15),
    panel.background = element_blank(),
    axis.title = element_text(size = 20),
    plot.margin = unit(c(0,0.8,0,0), "cm")
    )

ggsave(here::here("figures/mistake_cause_plot.jpg"), mistake_cause_plot, dpi = 300, height = 10, width = 14)

mistake_cause_plot
```

The number of excluded and missing responses.

```{r}
processed |> 
  filter(is.na(mistake_cause)) |>
  nrow()
```

# How much stress did you experience due to the retraction of the article?

```{r stressPlot, fig.path='../manuscript/figs/', dev=c('png', 'pdf'), fig.width=9, fig.height=6}
stress_plot_data <-
  processed |> 
  select(stress) |> 
  count(stress) |> 
  ungroup() |> 
  mutate(
    N = sum(n),
    percentage = round(n / N * 100, 0),
    stress = case_when(
      stress == 0 ~ "0\nNo stress at all",
      stress == 1 ~ "1",
      stress == 2 ~ "2",
      stress == 3 ~ "3",
      stress == 4 ~ "4",
      stress == 5 ~ "5",
      stress == 6 ~ "6\nExtreme stress"
      ),
    stress = factor(stress, levels = c("0\nNo stress at all" , "1" , "2" , "3" , "4" , "5" , "6\nExtreme stress"))
    )

y_max <- max(stress_plot_data$n) * 1.2

stress_plot <-
  stress_plot_data |> 
  ggplot() +
  aes(x = stress, y = n) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = paste0(percentage, "%")), size = 8, vjust = -0.3) +
  labs(x = paste0("How much stress did you experience due to the retraction of the article? (n =", unique(stress_plot_data$N), ")"),
       y = "Number of responses") +
  scale_y_continuous(expand = c(0,0), limit = c(0, y_max)) +
  theme(
    axis.ticks = element_line(color = "black"),
    axis.line = element_line(color = "black"),
    axis.text = element_text(color = "black", size = 13),
    panel.background = element_blank(),
    axis.title = element_text(size = 15),
    plot.margin = unit(c(0,0.8,0,0), "cm")
    )

ggsave(here::here("figures/stress_plot.jpg"), stress_plot, dpi = 300, height = 5, width = 10)

stress_plot
```

We will calculate the median and IQR for the question.

```{r}
# Median
median(processed$stress, na.rm = TRUE)

# Quantile 25%
quantile(processed$stress, .25, na.rm = TRUE)

# Quantile 75%
quantile(processed$stress, .75, na.rm = TRUE)
```

# Did you change anything in your research workflow because of the retraction?

```{r}
processed |> 
  select(workflow) |> 
  count(workflow) |> 
  ungroup() |> 
  mutate(
    N = sum(n),
    percentage = round(n / N * 100, 0)
    )
```

# Exploring the relationship between error causes and types

We are only looking at the most frequently occuring causes for errors.

```{r}
mistake_cause_type_plot_data <-
  processed |> 
  select(mistake_cause, mistake_type) |> 
  filter(!is.na(mistake_type)) |> 
  # Excluding missing and irrelevant responses
  filter(!is.na(mistake_cause)) |> 
  filter(mistake_cause %in% c("Inattention", "Technical issue", "Miscommunication", "Carelessness")) |> 
  count(mistake_cause, mistake_type) |> 
  group_by(mistake_cause) |>
  mutate(N = sum(n),
         percentage = round(n / N * 100, 0)) |> 
  arrange(mistake_cause)

# Number of all responses for this figure
mistake_cause_type_plot_data |> ungroup() |> summarise(all = sum(n))

mistake_cause_type_plot <-
  ggplot(data = mistake_cause_type_plot_data,
         aes(axis1 = mistake_cause, axis2 = mistake_type, y = n)) +
  geom_alluvium(aes(fill = mistake_cause)) +
  geom_stratum() +
  geom_text(stat = "stratum", aes(label = after_stat(stratum))) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(
    x = "Error causes and types",
    y = "Frequency"
  ) +
  theme_minimal() +
  theme(
    panel.grid = element_blank(),
    axis.line.y = element_line(color = "black"),
    axis.line.x = element_blank(),
    legend.position = "none",
    axis.text.x = element_blank(),
    axis.title = element_text(size = 15),
    axis.text.y = element_text(size = 12)
  )

ggsave(here::here("figures/mistake_cause_type_plot.jpg"), mistake_cause_type_plot, dpi = 300, height = 10, width = 14)

mistake_cause_type_plot
```
